Verification of concurrent design patterns with data

### 2019-02-26 2018-Arch-pNets/arch-pNets.tex
arch-pNets paper: revised introduction, related work and conclusion; run a spell checker.

\newcommand{\defn}[1]{Def.~\ref{defn:#1}}
\newcommand{\defntwo}[2]{Defs~\ref{defn:#1} and \ref{defn:#2}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\figs}[2]{Fig.~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\tab}[1]{Tab.~\ref{tab:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\res}[1]{(\ref{res:#1})}
\newcommand{\ex}[1]{Ex.~\ref{ex:#1}}
\newcommand{\exs}[2]{Ex.~\ref{ex:#1} and~\ref{ex:#2}}
\newcommand{\secn}[1]{Sect.~\ref{secn:#1}}
\newcommand{\app}[1]{App.~\ref{secn:#1}}
\newcommand{\rem}[1]{Note~\ref{rem:#1}}
\newcommand{\lem}[1]{Lem.~\ref{lem:#1}}
\newcommand{\cor}[1]{Cor.~\ref{cor:#1}}
\newcommand{\thm}[1]{Th.~\ref{thm:#1}}
\newcommand{\Thm}[1]{Theorem~\ref{thm:#1}}
\newcommand{\axs}[1]{Ax.~\ref{ax:#1}}
% \newcommand{\ax}[2]{\ref{ax:#1}.\ref{ax:#1:#2}}
\newcommand{\ax}[1]{Ax.~\ref{ax:#1}}
\newcommand{\prop}[1]{Prop.~\ref{prop:#1}}
\newcommand{\alg}[1]{Alg.~\ref{alg:#1}}
\newcommand{\asmp}[1]{Ass.~\ref{asmp:#1}}

@@

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.  Under
monotonicity assumptions on guards and assignments, composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
Additionally, the translation into  open pNet  gives
us a strong insight and a comparison of the coordination mechanisms
provided by the two frameworks.  The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

@@

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
BIP ensures that any property that holds on a BIP
model will also hold on the generated code.
The notion of BIP \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, adding
coordinator components and modifying the 
synchronisation patterns between components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie if two architectures guarantee two properties, their composition ensure the conjunction of the properties.
%
%several architectures are applied, each enforcing a safety
%property, the resulting system satisfies their conjunction.

This article goes one step further in the proof of properties and in the compositionality of architectures, but this step is a significant one: the compositional verification. To prove properties of BIP architectures
it is necessary to have a representation of the BIP architecture in a verifiable format. 
The verification problem has two unbounded parameters: 1) By nature, architectures have holes and are meant to interact with the interfaces of the component that will fill the hole; the properties must hold for all (well-typed) components that can be put inside the hole; 2) BIP interactions can transmit data, and properties might be dependent of the data, the domain of the data is generally huge or unbound and the value of transmitted data might have a significant impact on the properties.
Due to the nature of the problem, we propose to rely on a translation of BIP architectures into pNets, a direct verification of BIP could  be possible but pNets already have verification tools based on a SMT solver, and doing the same work for BIP would require  a large amount of similar development. More interestingly, relying on a translation to pNets allows us to compare the two models.

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems based on a parametrised and hierarchical model.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
It has been shown work~\cite{HMZ:PDP15} that pNets can  represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods, including various constructs and
languages for distributed objects. The VerCors  platform uses pNets to design
and verify  distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
pNets structure is static, but unbounded,  this allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can represent composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates value~\cite{HMZ-FORTE2016QBMZ-AVOCS18}. This encoding of open pNets into Z3 that is under development is the starting point of this article. We benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

We propose   to encode BIP architectures
into open pNets and to verify, using SMT techniques,
the properties of the architectures. The encoding also allows us to
compare BIP architectures and  pNet formalism.  
Overall, the encoding allows us to precisely compare the coordination
mechanisms of the two models but also the way component composition is
expressed in the two formalisms. 
%This also gives the opportunity
%to relate two  different visions on compositional verification.

@@

%\paragraph{Contributions}
The main contributions of this paper are: 1) The definition of a
theory of BIP architectures with data, including a theorem about
preservation of (data dependent) properties by 
compositions. 2) An encoding of architectures with data into  open pNets, allowing for analysis of their temporal properties using pNet's
software tools.
The  paper is illustrated by a running
example based on the failure monitor architecture from the CubETH
nanosatellite on-board software~\cite{CubETH-case-study}.

@@

%****************************************************************
We assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$, $e$ ranges over  expressions; and a set of \emph{Boolean
expressions} 
$\boolexprs{\variables}\subseteq\exprs{\variables}$, $g$ (guards) ranges over boolean expressions.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$.
We define $\assigns{\variables}$ as the set of variable assignments of the form: $(x_i := e_i)^{i \in I}$. 
%\[
%\assigns{\variables} \bydef{=}
%\Setdef{(x_i := e_i)^{i \in I}}{
%  x_i^{i \in I}\subseteq \variables,
%  e_i^{i \in I}\subseteq \exprs{\variables},
%  I \text{ is a finite index set}
%}
%\]
%to denote the set of \emph{variable assignments}. 
 $u$ ranges over sets of assignments.
%To be able to reason about the data flow between actions, we
%distinguish \emph{input variables} of the form $?x$ within terms; t
The function
$\vars(t)$ identifies the set of variables in a term.
%$t\in\Talg$, and $iv(t)$ returns its input variables.

@@

%\paragraph{Definitions.}
%\label{section:pNets}
A pLTS is a labelled transition system with variables that can be
used to characterise a state, as part of actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
Variables of each state are parwise disjoint.
We need to introduce input variables to reason about the data flow between actions.
We  define the set of parameterised actions a pLTS can use ($a$
ranges over action labels):\\
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
e_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $e_j^{j \in J}$ 
are expressions.

@@

%\paragraph{Operational Semantics.}
%\label{section:op-semantics}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNet is expressed as  an automaton where each transition coordinates the actions of several holes, 
the transition occurs if some predicates hold, and can involve  
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
	\label{defn:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes  and a set of states $\mathcal{S}$ is 
	a structure of the form:
\\[-2ex]	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{	\beta_j^{j\in J}, \Pred, \Post}
	{s \OTarrow {{\alpha}}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$. $\alpha$ is the resulting global action. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
	\assigns{\variables}$\ is 
	a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

@@

%% are \emph{coordinators}\mdash they are specified
%% completely, including the states and transitions defining their
%% \emph{behaviour}.
Component \emph{behaviour} is defined by states %% ($\{\cstate{0},
%% \cstate{1}, \cstate{2}\}$, for $\NameCtrl$; $\{\tstate{0}\}$, for
%% $\NameTimer$)
and transitions.
The initial states
\tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{t
\mapsto 0, z \mapsto \topinterval\}$
$\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
the incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$ where $\topinterval=(-\infty, +\infty)$.
%\footnote{%
%%
%  We use $\topinterval$ to denote the top interval in the inclusion
%  order, \ie $(-\infty, +\infty)$.
%%
%} 
The constants $\mathrm{Min}$ and $\mathrm{Max}$ are the parameters
of the architecture.

@@

%\addSB{%
%
\Thm{combining} implies that safe BIP systems can be designed
\emph{compositionally}.  Indeed, it is sufficient to verify that
1)~the applied architectures do enforce their characteristic
properties and 2)~that they are pairwise upwards compatible.  To
a large extent, item~2, \ie verification of the upwards
compatibility, can be carried out syntactically by analysing the
structure of the connectors defining the interaction models:
maximal progress is mostly relevant for connectors involving
triggers; whenever a trigger port is not common to both
architectures, upwards \emph{in}compatibility cannot arise.  The
next section is devoted to the encoding of architectures into
pNets, which addresses item~1 by symbolic verification.
%
%}

@@

\paragraph{Encoding the coordinators.}
The encoding of a coordinator $C$ is a pLTS with: 1)~an
 initial state and an   
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~an action algebra that matches the actions of the coordinator ports but adds, an additional boolean action parameter, and also, for each exported
variable $x$, a corresponding fresh input variable $?x'$ to allow updates during interactions, 3)~pLTS transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~additional loop transitions marked by
$\false$.
%% \noteLH{enleve le 3 et integre au 2}
Formally,  $\nopri[C] \bydef{=} \pNetTuple{S, s_0, \goesto[{\nopri[C]}]{}}$, such that
%
\begin{itemize}
\item $s_0 \not\in Q_C$ is fresh and $S = Q_C \cup \{s_0\}$,
  %% \noteEM{En principe les $vars(s)$ devraient etre disjoints, mais je
  %%   veux bien laisser ca sous le tapis tant qu'on n'implemente pas le codage...}
\item $\vars(s) = V_C \cup \setdef{?x'}{x \in \export[C](p), p \in a, s\goesto{a} }$, for all $s \in Q_C$, and $\vars(s_0) =
  \emptyset$,
%% \item the transition relation $\goesto[{\nopri[C]}]{}$ is defined as
%%   the union of three sets of transitions
%%   (with $\mathit{init} \bydef{=} \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$):
%% %
%%   \begin{multline}
%%     \label{eq:nopri:trans}
%%     {\goesto[{\nopri[C]}]{}} \bydef{=}
%%     \bsetdef{\tuple{s, a(\true), g, u, s'}}{
%%       \tuple{s, a, g, u, s'} \in \goesto{}}
%%     \\    
%%     \cup
%%     \bsetdef{\tuple{s, a(\false), \true, \noop, s}}{
%%       s \in Q_C, a \in P_C}
%%     %% \\
%%     \cup
%%     \bigl\{(s_0, \init, \true, \mathit{init}, q^0_C)\bigr\}
%%     \,.
%%   \end{multline}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{FailureTimerPNET-v4}
  \caption{The open pNET encoding the Failure Monitor architecture (\fig{schema:ArchFailure:BIP}) without the Max Progress priority model}
  \label{fig:FailureTimer:pNet}
%% %
%%   \noteSBin{The updated encoding exports more primed variables in the
%%     actions, \eg $tick(z,t,z',t')$ and $fail(zone, zone', b1)$ instead
%%     of $tick(z,t,z')$ and $fail(zone, b1)$, and SVs, \eg $\dots \land
%%     (t' = t) \land (zone' = zone)$ in the guard of SV0. Shall we
%%     update the figure to include these or, rather, explain that the
%%     figure is cleaned-out for readability?}
%
\end{figure}

@@

\section{Related work}
\label{secn:related}
Previous works on pNets and BIP architectures have been mentioned at the beginning of the paper. We  review here the closest related works outside the scope of pNets and BIP architectures, with a particular focus on compositional and parameterized verification.

@@

This can be sloved by the version with the Max Progress encoding.
This modifies both the pLTS transitions, and the synchronisation
vectors, as shown below:

>>>>

\newcommand{\defn}[1]{Def.~\ref{defn:#1}}
\newcommand{\defntwo}[2]{Defs~\ref{defn:#1} and \ref{defn:#2}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\figs}[2]{Fig.~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\tab}[1]{Tab.~\ref{tab:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\res}[1]{(\ref{res:#1})}
\newcommand{\ex}[1]{Ex.~\ref{ex:#1}}
\newcommand{\exs}[2]{Ex.~\ref{ex:#1} and~\ref{ex:#2}}
\newcommand{\secn}[1]{Sect.~\ref{secn:#1}}
\newcommand{\Secn}[1]{Section~\ref{secn:#1}}
\newcommand{\app}[1]{App.~\ref{secn:#1}}
\newcommand{\rem}[1]{Note~\ref{rem:#1}}
\newcommand{\lem}[1]{Lem.~\ref{lem:#1}}
\newcommand{\cor}[1]{Cor.~\ref{cor:#1}}
\newcommand{\thm}[1]{Th.~\ref{thm:#1}}
\newcommand{\Thm}[1]{Theorem~\ref{thm:#1}}
\newcommand{\axs}[1]{Ax.~\ref{ax:#1}}
% \newcommand{\ax}[2]{\ref{ax:#1}.\ref{ax:#1:#2}}
\newcommand{\ax}[1]{Ax.~\ref{ax:#1}}
\newcommand{\prop}[1]{Prop.~\ref{prop:#1}}
\newcommand{\alg}[1]{Alg.~\ref{alg:#1}}
\newcommand{\asmp}[1]{Ass.~\ref{asmp:#1}}

@@

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.  Under
monotonicity assumptions on guards and assignments, composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
%% Additionally, the translation into  open pNet  \noteEM{ICI}{gives
%% us a strong insight and a comparison of the coordination mechanisms
%% provided by the two frameworks}.
The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

@@

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
BIP ensures that any property that holds on a BIP
model will also hold on the generated code.
The notion of BIP \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, adding
coordinator components and modifying the 
synchronisation patterns between components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie if two architectures guarantee two properties, their composition ensures the conjunction of the properties.
%
%several architectures are applied, each enforcing a safety
%property, the resulting system satisfies their conjunction.

This article goes one step further in the proof of properties and in the compositionality of architectures, but this step is a significant one: the compositional verification. To prove properties of BIP architectures
it is necessary to have a representation of the BIP architecture in a verifiable format. 
The verification problem has two unbounded parameters: 1) By nature, architectures have holes and are meant to interact with the interfaces of the component that will fill the hole; the properties must hold for all (well-typed) components that can be put inside the hole; 2) BIP interactions can transmit data, and properties might be dependent of the data, the domain of the data is generally huge or unbounded and the values of transmitted data might have a significant impact on the properties.
%% Due to the nature of the problem, 
We propose to rely on a translation of BIP architectures into \emph{open pNets}.

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems based on a parameterised and hierarchical model.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
It has been shown in previous work~\cite{HMZ:PDP15} that pNets can  represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods, including various constructs and
languages for distributed objects. The VerCors  platform uses pNets to design
and verify  distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
The structure of pNets is static but unbounded.  This allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can represent composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates values~\cite{HMZ-FORTE2016QBMZ-AVOCS18}. The encoding of open pNets into Z3 that is under development is the starting point of this article. We benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

%% We propose   to encode BIP architectures
%% into open pNets and to verify, using SMT techniques,
%% the properties of the architectures. The encoding also allows us to
%% compare BIP architectures and  pNet formalism.  
%% Overall, the encoding allows us to precisely compare the coordination
%% mechanisms of the two models but also the way component composition is
%% expressed in the two formalisms.

%This also gives the opportunity
%to relate two  different visions on compositional verification.

@@

%\paragraph{Contributions}
The main contributions of this paper are: 1) The definition of a
theory of BIP architectures with data, including a theorem about
preservation of (data dependent) properties by 
compositions. 2) An encoding of architectures with data into  open pNets, allowing for analysis of their temporal properties using pNet's
software tools.  
The  paper is illustrated by a running
example based on the failure monitor architecture from the CubETH
nanosatellite on-board software~\cite{CubETH-case-study}.
%
This running example also relies on the \emph{maximal progress}
assumption, whereby larger interactions are preferred to smaller
ones.  Due to space limitations, we only discuss this
informally. However, proofs of the results provided in the
appendix formally account for maximal progress.

The rest of the paper is structured as follows.  In
\secn{preliminaries}, we present notations and background
material on pNets.  The theory of architectures with data is
presented in \secn{archs}.  In \secn{encoding}, we present the
encoding into open pNets and discuss verification of the running
example.  \Secn{related} discusses related work.
\Secn{conclusion} concludes the paper.

@@

%****************************************************************
We assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$, $e$ ranges over  expressions; and a set of \emph{Boolean
expressions} 
$\boolexprs{\variables}\subseteq\exprs{\variables}$, $g$ (guards) ranges over Boolean expressions.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$.
We define $\assigns{\variables}$ as the set of variable assignments of the form: $(x_i := e_i)^{i \in I}$. 
%\[
%\assigns{\variables} \bydef{=}
%\Setdef{(x_i := e_i)^{i \in I}}{
%  x_i^{i \in I}\subseteq \variables,
%  e_i^{i \in I}\subseteq \exprs{\variables},
%  I \text{ is a finite index set}
%}
%\]
%to denote the set of \emph{variable assignments}. 
 $u$ ranges over sets of assignments.
%To be able to reason about the data flow between actions, we
%distinguish \emph{input variables} of the form $?x$ within terms; t
The function
$\vars(t)$ identifies the set of variables in a term.
%$t\in\Talg$, and $iv(t)$ returns its input variables.

@@

%\paragraph{Definitions.}
%\label{section:pNets}
A pLTS is a labelled transition system with variables that can be
used to characterise a state, as part of actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
Variables of each state are pairwise disjoint.
We need to introduce input variables to reason about the data flow between actions.
We  define the set of parameterised actions a pLTS can use ($a$
ranges over action labels):
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
e_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $e_j^{j \in J}$ 
are expressions.

@@

%\paragraph{Operational Semantics.}
%\label{section:op-semantics}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNet is expressed as  an automaton where each transition coordinates the actions of several holes, 
the transition occurs if some predicates hold, and can involve  
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
  \label{defn:OpenTransitions}
  An \emph{open transition} over a
  set

  \noindent
  \begin{tabular}[b]{@{}p{0.86\columnwidth}c{}@{}}
    of holes $J$ and a set of states $\mathcal{S}$ is a structure
    of the form to the right,
%
    &
%
    %% \\[-2ex]	
    %% \begin{mathpar}
    \raisebox{1mm}[0pt]{
      $\inferrule*[myfraction=\reddottedrule]
      {	\beta_j^{j\in J}, \Pred, \Post}
      {s \OTarrow {{\alpha}}s'}$
      }
    %% \end{mathpar}
%
    \\
%
    \multicolumn{2}{p{\columnwidth}}{%
%
    where $s, s'\in\mathcal{S}$;
    $\beta_j\in\actions{\variables}$ is an action of the hole
    $j$; $\alpha$ is the resulting global action; {\Pred} is a
    predicate over the different variables of the terms, labels,
    and states $\beta_j$, $s$, $\alpha$; and $\Post\in
    \assigns{\variables}$ is a set of assignments that are the
    effects of the transition.  Open transitions are identified
    modulo logical equivalence on their predicate.
%
    }
  \end{tabular}
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

@@

%% are \emph{coordinators}\mdash they are specified
%% completely, including the states and transitions defining their
%% \emph{behaviour}.
Component \emph{behaviour} is defined by states %% ($\{\cstate{0},
%% \cstate{1}, \cstate{2}\}$, for $\NameCtrl$; $\{\tstate{0}\}$, for
%% $\NameTimer$)
and transitions.
The initial states
\tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{t
\mapsto 0, z \mapsto \topinterval\}$,
$\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
the incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$ where $\topinterval=(-\infty, +\infty)$.
%\footnote{%
%%
%  We use $\topinterval$ to denote the top interval in the inclusion
%  order, \ie $(-\infty, +\infty)$.
%%
%} 
The constants $\mathrm{Min}$ and $\mathrm{Max}$ are the parameters
of the architecture.

@@

%\addSB{%
%
\Thm{combining} implies that safe BIP systems can be designed
\emph{compositionally}:  it is sufficient to verify that
1)~the applied architectures do enforce their characteristic
properties and 2)~they are pairwise upwards compatible.  To
a large extent, %% item~2, \ie verification of the upwards
%% compatibility,
the latter
can be carried out syntactically by analysing the
structure of the connectors that define the interaction models.
%% maximal progress is mostly relevant for connectors involving
%% triggers; whenever a trigger port is not common to both
%% architectures, upwards \emph{in}compatibility cannot arise.
The
next section is devoted to the encoding of architectures into
pNets, addressing item~1 by symbolic verification.
%
%}

@@

\paragraph{Encoding the coordinators.}
The encoding of a coordinator $C$ is a pLTS with: 1)~an
 initial state and an   
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~an action algebra that matches the actions of the coordinator ports but adds, an additional Boolean action parameter, and also, for each exported
variable $x$, a corresponding fresh input variable $?x'$ to allow updates during interactions, 3)~pLTS transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~additional loop transitions marked by
$\false$.
%% \noteLH{enleve le 3 et integre au 2}
Formally,  $\nopri[C] \bydef{=} \pNetTuple{S, s_0, \goesto[{\nopri[C]}]{}}$, such that
%
\begin{itemize}
\item $s_0 \not\in Q_C$ is fresh and $S = Q_C \cup \{s_0\}$,
  %% \noteEM{En principe les $vars(s)$ devraient etre disjoints, mais je
  %%   veux bien laisser ca sous le tapis tant qu'on n'implemente pas le codage...}
\item $\vars(s) = V_C \cup \setdef{?x'}{x \in \export[C](p), p \in a, s\goesto{a} }$, for all $s \in Q_C$, and $\vars(s_0) =
  \emptyset$,
%% \item the transition relation $\goesto[{\nopri[C]}]{}$ is defined as
%%   the union of three sets of transitions
%%   (with $\mathit{init} \bydef{=} \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$):
%% %
%%   \begin{multline}
%%     \label{eq:nopri:trans}
%%     {\goesto[{\nopri[C]}]{}} \bydef{=}
%%     \bsetdef{\tuple{s, a(\true), g, u, s'}}{
%%       \tuple{s, a, g, u, s'} \in \goesto{}}
%%     \\    
%%     \cup
%%     \bsetdef{\tuple{s, a(\false), \true, \noop, s}}{
%%       s \in Q_C, a \in P_C}
%%     %% \\
%%     \cup
%%     \bigl\{(s_0, \init, \true, \mathit{init}, q^0_C)\bigr\}
%%     \,.
%%   \end{multline}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{FailureTimerPNET-v4}
  \caption{The open pNet encoding the Failure Monitor architecture (\fig{schema:ArchFailure:BIP}) without the Max Progress priority model}
  \label{fig:FailureTimer:pNet}
%% %
%%   \noteSBin{The updated encoding exports more primed variables in the
%%     actions, \eg $tick(z,t,z',t')$ and $fail(zone, zone', b1)$ instead
%%     of $tick(z,t,z')$ and $fail(zone, b1)$, and SVs, \eg $\dots \land
%%     (t' = t) \land (zone' = zone)$ in the guard of SV0. Shall we
%%     update the figure to include these or, rather, explain that the
%%     figure is cleaned-out for readability?}
%
\end{figure}

@@

\section{Related work}
\label{secn:related}
%% Aside from previous work on pNets and BIP architectures mentioned in \secn{introduction}. We  review here the closest related works outside the scope of pNets and BIP architectures, with a particular focus on compositional and parameterized verification.

@@

This can be solved by the version with the Max Progress encoding.
This modifies both the pLTS transitions, and the synchronisation
vectors, as shown below:
